Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-06-15 omrp_lb_rev.tex
Review of theoretical efficiency gains and comparison to NV

xxx Add new graphs here \bigskip

xxx Write new text, etc. \bigskip

xxx Revise the recommendations on gammabar \bigskip

Because variance reduction drops quickly as $\gammab$ drops below $1$, we recommend using an intermediate value of $\gammab$, such as $1/3$ or $1/4$, to gain the majority of the variance reduction, while reducing the number of optimization problems to be solved. 
Warm-starting the algorithm to solve the sampling problems, when such a scheme is available, can significantly reduce the computation times.

>>>>

\begin{figure}[htb!]
	\centering
	\subfloat[][]{
	   \label{fig:efficiency_theory}
	   \includegraphics[width=.5\textwidth]{images/comp_efficiency_theory}}
	\subfloat[][]{
	   \label{fig:efficiency_nv}
	   \includegraphics[width=.5\textwidth]{images/nv_comp_efficiency}}
		\caption{
		Theoretical computational efficiency results for overlapping batches (left) and a comparison with measured computational efficiency for the Newsvendor problem using insertion sort (right).
		Both graphs are scaled so that nonoverlapping batches produces a computational efficiency of 1.
		}
\label{fig:efficiency}
\end{figure}

The increased computational workload from solving overlapping batches raises the question of the computational efficiency of this method to estimate $VG$.
In this setting, computational efficiency may be defined as $[T(\gammab)\var{VG}]$ where $T(\gammab)$ is the computational time.
Because variance reduction is known from (\ref{eq:var_reduct_formula}) we need only examine the computational time, which can be decomposed as $T(\gammab) = t(\gammab)\left(\nb\right)$ where $t(\gammab)$ is the average computation time per batch.
Warm-starting an algorithm may be used to decrease the batch solution time for $\gammab < 1$, leading to a possible increase in computational efficiency.

To investigate this, we assumed that computation time per batch can be approximated as $t(\gammab) \propto \gammab^p$ for some $p \geq 0$.
Results for some values of $p$ are shown in Figure \ref{fig:efficiency_theory}.
Efficiency gains increase with larger exponent $p$, but are positive even for small values of $p$ when $\gammab$ is sufficiently close to 1.
At $p=1$, i.e., for problems where computational time is proportional to the number of samples replaced in the batch, computational efficiency is maximized with maximally overlapping batches.

We tested this theoretical relationship using the Newsvendor problem solved with insertion sort.
The average computation time per batch was found empirically to be well approximated by power law with $p = 0.71$.
The measured computational efficiency is compared to the theoretical result in Figure \ref{fig:efficiency_nv}.

In cases where the exponent $p$ can be estimated the above theoretical method may be used to determine a reasonable value for $\gammab$.
We can see in Figure \ref{fig:efficiency} that for some values of $p$ efficiency gian is nearly flat over a region of $\gammab$.
In this case we recommend choosing a $\gammab$ near the right side of this region.
This choice keeps efficiency high while decreasing computation time and preventing large decreases in efficiency in the event that $p$ was overestimated.
